Towards Performance-portable Parallel Programming in Computer Science Education

### 2020-01-22 src/2.Kokkos.tex
Minor changes.

Both examples provide the semantic information on what to parallelize, however, they do not provide information on how to map the parallel execution on modern parallel architectures. In order to define what further information is needed for a generic support on concurrency on modern computer architectures, it is important to define an abstract machine model first.

@@

An instance of a memory space provides a concrete method for the application programmer to request data storage allocations. Following the architecture as discussed earlier, the multi-core processor may contain multiple memory spaces that abstract on-package memory, slower DRAM and non-volatile memories. Accelerators can provide an additional memory space through its local on-package memory. The programmer is free to decide where each data structure is allocated by requesting the corresponding memory space by instantiating the according memory space. Kokkos provides the appropriate abstraction of memory allocation and data management. We believe that an abstract representation of execution- and memory devices is a key property towards performance-portable parallel programming.

@@

Kokkos defines a pure interface for C++ and uses C++ as the base programming language. This has multiple reasons. It is our experience that application developers of scientific and high-performance computing applications at Sandia National Laboratories express the demand for C++ and a generic parallel programming model to support their codes. Following their estimates, porting efforts towards adapting to vendors, programming models, APIs and software releases represents a burden worth addressing. Further, the possibility to maintain pure C++ codes is appealing from the code development and debugging perspective. 
Lastly, from the programming model perspective, C++ offers template metaprogramming which is well suited to implement generic APIs and libraries. In this case, class specialization and templating allow for compile-type generated types and optimizations for a given hardware architecture. It is worth to mention that Kokkos avoids the use of pragma annotations for the sake of treating concurrency as a first-class concept in C++.

>>>>

Both examples provide the semantic information on what to parallelize, however, they do not provide information on how to map the parallel execution on modern parallel architectures. In order to define what further information is needed to express concurrency on modern computer architectures, it is important to define an abstract machine model first.

@@

An instance of a memory space provides a concrete method for the application programmer to request data storage allocations. Following the architecture as discussed earlier, the multi-core processor may contain multiple memory spaces that abstract on-package memory, slower DRAM and non-volatile memories. Accelerators can provide an additional memory space through its local on-package memory. An architectural layout is shown in Figure~\ref{fig:memspace}. The programmer is free to decide where each data structure is allocated by requesting the corresponding memory space. Programmatically this is achieved by instantiating the according memory space. Kokkos provides the appropriate abstraction of memory allocation and data management. We believe that an abstract representation of execution- and memory devices is a key property towards performance-portable parallel programming.

@@

Kokkos defines a pure interface for C++ and uses C++ as a programming language. This has multiple reasons. It is our experience that application developers of scientific and high-performance computing applications at Sandia National Laboratories express the demand for C++ and a generic parallel programming model to support their codes. Following their estimates, porting efforts towards adapting to vendors, programming models, APIs and software releases represents a burden worth addressing. Further, the possibility to maintain pure C++ codes is appealing from the code development and debugging perspective. 
Lastly, from the programming model perspective, C++ offers template metaprogramming which is well suited to implement generic APIs and libraries. In this case, class specialization and templating allow for compile-type generated types and optimizations for a given hardware architecture. It is worth to mention that Kokkos avoids the use of pragma annotations for the sake of treating concurrency as a first-class concept in C++.
